Automatic Amortized Analysis for Imperative Code

### 2014-03-07 main.tex
use explicit pairs for resource monoid
add soundness part 1

This big step semantics counts the number of iterations of
while loops in a program as it executes it.  Non-deterministically,
the {\sc S:Abort} rule can be called to interrupt a computation.
It has the same effect as raising an exception.  We symbolize a
frozen heap with the notation $\circ H$, this kind of heap just
bubbles up through the semantics. In the rules, $X$ stands for
either $\circ$ or $H$.

@@

\and \Rule{S:WhileStep}
{  k < \llbracket x - y \rrbracket_H
\\ (H, S) \Downarrow_{q_1} H'
\\ (H', \code{while}(k < x - y)~S) \Downarrow_{q_2} X
}
{ (H, \code{while}(k < x - y)~S) \Downarrow_{M_w^1 \cdot q_1 \cdot M_w^2 \cdot q_2} X }

\and \Rule{S:WhileStepAbort}
{  k < \llbracket x - y \rrbracket_H
\\ (H, S) \Downarrow_{q_1} \circ
}
{ (H, \code{while}(k < x - y)~S) \Downarrow_{M_w^1 \cdot q_1 } \circ }

\and \Rule{S:If}
{  (H, S_i) \Downarrow_q X
\\ i \in \{1, 2\}
}
{ (H, \code{if}(?)~S_1~\code{else}~S_2) \Downarrow_{M_c^i \cdot q} X }

\and \Rule{S:Seq}
{  (H, S_1) \Downarrow_{q_1} H'
\\ (H', S_2) \Downarrow_{q_2} X
}
{ (H, S_1; S_2) \Downarrow_{M_s^1 \cdot q_1 \cdot M_s^2 \cdot q_2} X }

\and \Rule{S:SeqAbort}
{ (H, S_1) \Downarrow_q \circ }
{ (H, S_1; S_2) \Downarrow_q \circ }

@@

>>>>

This big step semantics counts the number of iterations of
while loops in a program as it executes it.  Non-deterministically,
the {\sc S:Abort} rule can be called to interrupt a computation.
It has the same effect as raising an exception.  We symbolize a
frozen computation with the notation $\circ$.  In the rules, $X$
stands for either $\circ$ or $H$.

@@

\and \Rule{S:WhileStep}
{  k < \llbracket x - y \rrbracket_H
\\ (H, S) \Downarrow_{(q_1, q'_1)} H'
\\ (H', \code{while}(k < x - y)~S) \Downarrow_{(q_2, q'_2)} X
}
{ (H, \code{while}(k < x - y)~S) \Downarrow_{M_w^1 \cdot (q_1, q'_1) \cdot M_w^2 \cdot (q_2, q'_2)} X }

\and \Rule{S:WhileStepAbort}
{  k < \llbracket x - y \rrbracket_H
\\ (H, S) \Downarrow_{(q_1, q'_1)} \circ
}
{ (H, \code{while}(k < x - y)~S) \Downarrow_{M_w^1 \cdot (q_1, q'_1) } \circ }

\and \Rule{S:If}
{  (H, S_i) \Downarrow_{(q, q')} X
\\ i \in \{1, 2\}
}
{ (H, \code{if}(?)~S_1~\code{else}~S_2) \Downarrow_{M_c^i \cdot (q, q')} X }

\and \Rule{S:Seq}
{  (H, S_1) \Downarrow_{(q_1, q'_1)} H'
\\ (H', S_2) \Downarrow_{(q_2, q'_2)} X
}
{ (H, S_1; S_2) \Downarrow_{M_s^1 \cdot (q_1, q'_1) \cdot M_s^2 \cdot (q_2, q'_2)} X }

\and \Rule{S:SeqAbort}
{ (H, S_1) \Downarrow_{(q,q')} \circ }
{ (H, S_1; S_2) \Downarrow_{(q,q')} \circ }

@@

\section{Soundness theorem}

\begin{theorem}[Terminating case]
For any $S$, $H$, $H'$, $q$, $q'$, $\Gamma$, $\Gamma'$, $Q$ and $Q'$ such
that $\Gamma; Q \vdash S \dashv Q'; \Gamma'$ and $H \models \Gamma$ and
$(H,s) \Downarrow_{(q,q')} H'$, we have
$H' \models \Gamma'$ and
$q \le \Phi_H(Q)$ and
$\Phi_H(Q) \ge q' + \Phi_{H'}(Q')$.
\end{theorem}
